\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Changes to eval

Using \sys's standard library of annotations for POSIX and GNU commands~\sx{parallelizability}, the vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 commands outside these classes ($<3\%$) needed a single-record annotation.

@@

\heading{Parallelism}
%
% All experiments are run on the same machine and 
% \nv{Eurosys: This /\ is described above}
The level of parallelism is configured using \sys's \ttt{--parallel=N}, a flag that determines the number of times that \sys applies its transformations---determining the ``width'' of the parallel script.
The value  of \ttt{N} is shown as the x-axis of Fig.\ref{fig:microbenchmarks}, called \emph{parallelism} throughout this section.
We cannot control the initial parallelism of a script (\eg a command could spawn 10 processes), and thus parallel scripts produced by \sys often reach maximum parallelization benefits with a value of \ttt{N} that is smaller than the physical cores of the hardware setup (in our case 64).

@@

We evaluate \sys under multiple configurations on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.
% \nv{Eurosys: goal of experiment: decouple compiler from runtime, and talk about all overheads}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for 2--64$\times$-parallelism (\S\ref{eval}, Parallelism).}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
      % (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      % Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

\heading{Programs}
%>\kk{How can we include a result for grep-light?  This is the IO
%>  intensive CPU-light grep. \sys has no speedup but also no slowdown
%>  (even if it IO heavy). Should we just mention that in text in
%>  passing?}
%>
%>\kk{Remember to mention that for spell and bigrams we dont show other
%>  parallel because they are all close to 1 (no speedup).}
Tab.~\ref{tab:eval} summarizes the first collection of programs.
%  shown in Fig.~\ref{fig:microbenchmarks}, which in turn reports on their performance characteristics.
Grep and Grep-light are in \sta, with the former centered around an expensive DFA-based backtracking expression.
% a DFA-based regular expression matching phase;
 % while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
Sort is a short script centered around a \pur command.
Wf and Top-n are based on McIlroy's classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
Spell, based on the original \ttt{spell} developed by Johnson~\cite{bentley1985spelling}, is another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in a dictionary.
Shortest-scripts extracts the 15 shortest scripts in the user's \ttt{PATH}, using the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}~\cite[pg. 7]{taylor2004wicked}.
Diff and Set-diff compare streams via a \ttt{diff} (in \npu, non-parallelizable) and \ttt{comm} (in \pur), respectively.
Sort-sort uses consecutive \pur commands without interleaving them with commands that condense their input size (\eg \ttt{uniq}).
Finally, the two Bi-grams replicate and shift a stream by one entry to calculate bigrams, but Bi-gram-opt's shift, \ttt{sort}, and \ttt{uniq} are merged to one command to use a more efficient aggregator.

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}

% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
It also contains Filter and Bigrams-opt (not in Fig.~\ref{fig:microbenchmarks}).
Filter is used to evaluate \sys's impact on IO-intensive scripts (\sys achieves a 1.5--2.5$\times$ speedup).
Bigrams-opt is an optimized version of Bigrams that groups 5 commands into one \pur command and thus has a more efficient $map$ and $aggregate$ implementation.
\sys achieves up to $30\times$ speedup for this script (significantly higher than what is achieved for the standard Bigrams Fig.\ref{fig:microbenchmarks}).
% \nv{Eurosys: I think this should be part of a "do custom aggregators help?" question, not the main results}.
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
%
\todo{These results explain a detail before getting the main points across: A is better than B, C is worse except G}
As discussed earlier (\S\ref{eval}, Parallelism), \sys does not have control over the input script's parallelism.
Thus, it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system.
This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}.
As described in Fig.\ref{optimizer}, these eager implementations perform worse than \sys's optimized \ttt{eager} relay in all cases (except for a negligible difference in Shortest-Scripts).
A \ttt{split} benefits scripts with \tpur or \tnpu commands, but does not negatively affect the rest.

>>>>

Using \sys's standard library of annotations for POSIX and GNU commands~\sx{parallelizability}, the vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6  ($<3\%$) commands, outside this library, needed a single-record annotation.

@@

\heading{Parallelism}
The level of parallelism is configured using \sys's \ttt{--width=N}, a flag that determines the number of times that \sys applies its transformations and the resulting width of the data-parallel DFG.
% The value  of \ttt{N} is shown as the x-axis of Fig.\ref{fig:microbenchmarks}. 
We cannot control the initial parallelism of a script (\eg a command could spawn 10 processes), and thus parallel scripts produced by \sys often reach maximum parallelization benefits with a value of \ttt{N} that is smaller than the physical cores of the hardware setup (in our case 64).

@@

We evaluate \sys on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.
The goal is to evaluate
  (i) the performance benefits stemming from \sys's  DFG transformations alone, including how \ttt{--width} affects speedup, 
  (ii) performance benefits stemming from \sys's runtime primitives, showing results for several configurations of the runtime lattice (Fig.~\ref{fig:lattice}).

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for \ttt{width=} 2--64.}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
  For some pairs of configurations, \sys produces identical parallel scripts and thus only one is shown.
%   Two types of pairs where x y z coincide
%   - \sys adds no split, whenever whenever the DFG width is constant for the entire pipeline. In these cases the blue line is identical to the red line, which not shown.
%    Conversely, whenever both blue and red lines are shown, split nodes are needed due to changes in the pipeline width and result in performance improvements.
% 
%   - without split, the script is sequential (spell, bigrams)
% Early on they have a pure or no parallelizable pure
% and thus introduce an aggregator resulting in width of 1 right after, and requires a split.
% Thus, any configuration that does not has \ttt{split} is practically sequential beyond that point---\ie No Eager, Blocking Eager, and \sys w/o Split see no speedup.
% 
% Blocking eager: data parallelism, but no task parallelism
% No eager: blocking FIFO does not allow pushing -> underutilize CPU
% 
% The reason why Blocking Eager is usually slower than No Eager is due to \ttt{agg}regators that read multiple input streams at the same time---such \ttt{sort -m}.
% 
% Sort-sort possibly shows the full spectrum:
%   (iii) \sys's \ttt{eager} offers benefits despite the lack of \ttt{split} because it fully parallelizes the first \ttt{sort} while maintaining task parallelism
%   (iv) \sys gets full benefits because \ttt{split}ting allows parallelizing the second \ttt{sort} too.

  % Configurations where the parallel script produced is identical to an existing configuration are not shown.

      % (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      % Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

\heading{Programs}
%>\kk{How can we include a result for grep-light?  This is the IO
%>  intensive CPU-light grep. \sys has no speedup but also no slowdown
%>  (even if it IO heavy). Should we just mention that in text in
%>  passing?}
%>
%>\kk{Remember to mention that for spell and bigrams we dont show other
%>  parallel because they are all close to 1 (no speedup).}
Tab.~\ref{tab:eval} summarizes the first collection of programs.
%  shown in Fig.~\ref{fig:microbenchmarks}, which in turn reports on their performance characteristics.
NFA-Regex and Filter are in \sta, with the former centered around an expensive \todo{NFA-based} backtracking expression and the latter being an IO-intensive script.
% a DFA-based regular expression matching phase;
 % while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
Sort is a short script centered around a \pur command.
Wf and Top-n are based on McIlroy's classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
Spell, based on the original \ttt{spell} developed by Johnson~\cite{bentley1985spelling}, is another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in a dictionary.
Shortest-scripts extracts the 15 shortest scripts in the user's \ttt{PATH}, using the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}~\cite[pg. 7]{taylor2004wicked}.
Diff and Set-diff compare streams via a \ttt{diff} (in \npu, non-parallelizable) and \ttt{comm} (in \pur), respectively.
Sort-sort uses consecutive \pur commands without interleaving them with commands that condense their input size (\eg \ttt{uniq}).
Finally, the two Bi-grams replicate and shift a stream by one entry to calculate bigrams, but Bi-gram-opt's groups 5 commands into one \pur command that has a more efficient $map$ and $aggregate$ implementation.

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of \ttt{width=}2--64.
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \ttt{width=}\{2, 4, 8, 16, 32, 64\} are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
Average speedups for No Eager, \ie only \sys' transformations without its runtime support, are 1.63, 2.54, 3.86, 5.93 7.46 9.35$\times$.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
Filter and Bigrams-opt are not in Fig.~\ref{fig:microbenchmarks}, for which \sys achieves a speedup 1.5--2.5$\times$ and $30\times$.
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
% TODO: Add
% * for scripts only with commands in \sta, we get linear speedup.
% * eager is better than no eager and blocking eager for all scripts
% * A \ttt{split} benefits scripts with \tpur or \tnpu commands, but does not negatively affect the rest.
% * \ttt{width}
% * why blocking is less than no eager
% ( standard vs non-standard Bigrams Fig.\ref{fig:microbenchmarks}).
\todo{These results explain a detail before getting the main points across: A is better than B, C is worse except G}
As discussed earlier (\S\ref{eval}, Parallelism), \sys does not have control over the input script's parallelism.
Thus, it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system.
This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}.
As described in Fig.\ref{optimizer}, these eager implementations perform worse than \sys's optimized \ttt{eager} relay in all cases (except for a negligible difference in Shortest-Scripts).
